---
title: "Tidy Tuesday - Astronaut Missions"
author: "Matthew Wilson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  fig.height = 5,
  fig.width = 8,
  paged.print=FALSE,
  warning = TRUE,
  message = FALSE)

```

#### Necessary package and themes

```{r}

library(tidyverse)
library(tidymodels)
library(skimr)
library(janitor)
library(corrr)
library(vip)
library(visdat)
library(reticulate)

pal <- c("#0073C2FF","#EFC000FF","#868686FF","#CD534CFF","#7AA6DCFF",
         "#003C67FF","#8F7700FF","#3B3B3BFF","#A73030FF","#4A6990FF")

theme_set(theme_minimal())
theme_update(axis.line.x = element_line(color="dark grey"),
             axis.line.y = element_line(color="dark grey"))



```

### Summary of objective

Building boostrap models using a #TidayTuesday dataset with astronaut missions.  

**Predict** `duration` in space

For this analysis, we are going to build a bagging, i.e. bootstrap aggregating, model. This is an ensembling and model averaging method that:  

 -- improves accuracy and stability
 -- reduces overfitting and variance  

In tidymodels, you can create bagging ensemble models with baguette, a parsnip-adjacent package. The baguette functions create new bootstrap training sets by sampling with replacement and then fit a model to each new training set. These models are combined by averaging the predictions for the regression case, like what we have here (by voting, for classification).

#### Loading the dataset
```{r}
astronauts <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-14/astronauts.csv")

astronauts %>%
  count(in_orbit, sort = TRUE)

```

```{r}

astronauts %>%
  mutate(
    year_of_mission = 10 * (year_of_mission %/% 10),
    year_of_mission = factor(year_of_mission)) %>%
  ggplot(aes(year_of_mission, hours_mission,
    fill = year_of_mission, color = year_of_mission)) +
  geom_boxplot(alpha = 0.2, size = 1.5, show.legend = FALSE) +
  scale_y_log10() +
  scale_color_manual(values = pal)+ scale_fill_manual(values = pal)+
  labs(x = NULL, y = "Duration of mission in hours")

```

```{r}

astronauts_df <- astronauts %>%
  select(
    name, mission_title, hours_mission,
    military_civilian, occupation, year_of_mission, in_orbit
  ) %>%
  mutate(
    in_orbit = case_when(
      str_detect(in_orbit, "^Salyut") ~ "Salyut",
      str_detect(in_orbit, "^STS") ~ "STS",
      TRUE ~ in_orbit
    ),
    occupation = str_to_lower(occupation)
  ) %>%
  filter(hours_mission > 0) %>%
  mutate(hours_mission = log(hours_mission)) %>%
  na.omit()

ggplot(astronauts_df, aes(hours_mission))+
  geom_histogram()
```

#### Split into test and train datasets

```{r}

data_split <- initial_split(astronauts_df, strata = hours_mission)
X_train <- training(data_split)
X_test <- testing(data_split)

```

#### Preprocessing steps

```{r}

astro_recipe <- 
  recipe(hours_mission ~ ., data = X_train) %>%
  update_role(name, mission_title, new_role = 'id') %>% 
  step_other(occupation, in_orbit,
      threshold = 0.005, other = 'Other') %>%
  step_dummy(all_nominal(), -has_role('id'))

```


```{r}

astro_recipe %>% 
  prep() %>% 
  juice() %>% 
  names()

```

#### Workflow  

This allows us to pass process and prep in one step

```{r}

boost_wf <-
  workflow() %>% 
  add_recipe(astro_recipe)

```

#### Model Specifications

```{r}

library(baguette)

tree_spec <-
  bag_tree() %>% 
  set_engine('rpart', times = 25) %>% 
  set_mode('regression')

tree_spec

```

```{r}

mars_spec <-
  bag_mars() %>% 
  set_engine('earth', times = 25) %>% 
  set_mode('regression')

mars_spec

```

#### Fit the models

```{r}

tree_fit <- 
  boost_wf %>% 
  add_model(tree_spec) %>% 
  fit(X_train)

```

```{r}

mars_fit <-
  boost_wf %>% 
  add_model(mars_spec) %>% 
  fit(X_train)

```

```{r}
tree_fit
```

```{r}
mars_fit
```

#### Evaluate model resutls


```{r}

test_results <-
  X_test %>% 
  bind_cols(predict(tree_fit, X_test)) %>% 
  rename(.pred_tree = .pred) %>% 
  bind_cols(predict(mars_fit, X_test)) %>% 
  rename(.pred_mars = .pred)

test_results %>% metrics(hours_mission, .pred_tree)
test_results %>% metrics(hours_mission, .pred_mars)

```



#### Create dummy new prediction data

```{r}

new_astronauts <- crossing(
  in_orbit = fct_inorder(c("ISS", "STS", "Mir", "Other")),
  military_civilian = "civilian",
  occupation = "Other",
  year_of_mission = seq(1960, 2020, by = 10),
  name = "id", mission_title = "id"
) %>%
  filter(
    !(in_orbit == "ISS" & year_of_mission < 2000),
    !(in_orbit == "Mir" & year_of_mission < 1990),
    !(in_orbit == "STS" & year_of_mission > 2010),
    !(in_orbit == "STS" & year_of_mission < 1980)
  )

new_astronauts
```

#### Predict on new data set

```{r}
new_astronauts %>%
  bind_cols(predict(tree_fit, new_astronauts)) %>%
  ggplot(aes(year_of_mission, .pred, color = in_orbit)) +
  geom_line(size = 1.5, alpha = 0.7) +
  geom_point(size = 2) +
  scale_color_manual(values = pal)+
  labs(
    x = NULL, y = "Duration of mission in hours (predicted, on log scale)",
    color = NULL, title = "How did the duration of astronauts' missions change over time?",
    subtitle = "Predicted using bagged decision tree model"
  )
```

```{r}
new_astronauts %>%
  bind_cols(predict(mars_fit, new_astronauts)) %>%
  ggplot(aes(year_of_mission, .pred, color = in_orbit)) +
  geom_line(size = 1.5, alpha = 0.7) +
  geom_point(size = 2) +
  scale_color_manual(values = pal)+
  labs(
    x = NULL, y = "Duration of mission in hours (predicted, on log scale)",
    color = NULL, title = "How did the duration of astronauts' missions change over time?",
    subtitle = "Predicted using bagged MARS model"
  )
```

#### Test random forest

```{r}

rf_spec <-
  rand_forest(mtry = 10,
              min_n = 1,
              trees = 1000) %>% 
  set_mode('regression') %>% 
  set_engine('randomForest')

rf_fit <-
  boost_wf %>% 
  add_model(rf_spec) %>% 
  fit(X_train)

  rf_fit %>% 
  predict(X_test) %>% 
  bind_cols(X_test) %>% 
  metrics(hours_mission,.pred)

```

```{r}
new_astronauts %>%
  bind_cols(predict(rf_fit, new_astronauts)) %>%
  ggplot(aes(year_of_mission, .pred, color = in_orbit)) +
  geom_line(size = 1.5, alpha = 0.7) +
  geom_point(size = 2) +
  scale_color_manual(values = pal)+
  labs(
    x = NULL, y = "Duration of mission in hours (predicted, on log scale)",
    color = NULL, title = "How did the duration of astronauts' missions change over time?",
    subtitle = "Predicted using random forest model"
  )
```


#### Test an svm

```{r}

svm_spec <-
  svm_poly(cost = 10, scale_factor = 0.1) %>%  
  set_engine('kernlab') %>% 
  set_mode('regression') 

svm_fit <-
  boost_wf %>% 
  add_model(svm_spec) %>% 
  fit(X_train)

  svm_fit %>% 
  predict(X_test) %>% 
  bind_cols(X_test) %>% 
  metrics(hours_mission,.pred)
  
```


```{r}

new_astronauts %>%
  bind_cols(predict(svm_fit, new_astronauts)) %>%
  ggplot(aes(year_of_mission, .pred, color = in_orbit)) +
  geom_line(size = 1.5, alpha = 0.7) +
  geom_point(size = 2) +
  scale_color_manual(values = pal)+
  labs(
    x = NULL, y = "Duration of mission in hours (predicted, on log scale)",
    color = NULL, title = "How did the duration of astronauts' missions change over time?",
    subtitle = "Predicted using polynomial svm model"
  )
```

